---
title: "API 222 Problem Set 1"
author: "Liz Masten"
date: "9/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(AER)
library(tidyverse)

```

# Conceptual Questions 

### 1. For each of the following questions, state: (6 pts)
##### (1) Whether it is a regression question or a classification question
##### (2) Whether we are interested in inference or prediction
     
(a) The New York City Mayor’s Office plans to solicit construction bids for a school renovation project. They want to know if each bid will be finished on time. They have past
data on NYC construction bids, including information on project type, construction
company characteristics, budget estimates and whether the bid was finished on time.

This question is a classification question because the output variable ("finished on time") is discrete. This is discrete because the projects will be either finished on time or they won't be. We are interested in prediction here because we are using existing data (the past data on NYC construction bids) to predict outcomes revolving around new information (the new construction bids). 

(b) The mayors office also care about the budget considerations. Officials want to avoid
projects that understate their true budget and pick bids that will have final spending
close to the proposed budget. The vast majority of projects are over budget, so the
office wants to know to how much over budget each potential bid will be.

This question is a regression question because the outcome variable (how much over budget) is continuous. We are interested in prediction again because we are still using data on past construction bids to predict by how much each new bid will be over budget. 

(c) The mayor noticed that local contractors seemed to win more construction bids. The
mayor was interested in if local contractors are better at assessing the needs of the
community or if political connections were driving this trend. The mayor decided to
implement a blind submission process where the location and name of the firm are hidden for two years. He wants you to analyze whether blind hiring changes the likelihood
of a local bid being chosen. For each proposal for the last 10 years, you have details on
year submitted, location, number of employees, firm age, number of previous contracts,
total portfolio amount, whether the proposal was accepted, etc.

This question is a classification question because the output variables are categorical or discrete. Here, we are interested in inference because we are using existing data to look at the effects within that data.  

### 2. Flexible models versus inflexible models (1 pt)

(a) I have two models, one with high bias and variance and one with low bias and low
variance. I should choose the model with high bias. True or False?

FALSE - ideally we should have low bias and low variance. 

(b) KNN and linear regression are both parametric models, as they both have decision rules.
True or False?

FALSE - linear regression is parametric but KNN is not. 

(c) A second order polynomial will always have lower bias than a linear model. True or
False?

TRUE - but it might have a higher variance due to overfitting, depending on the data. 


(d) We should run our model multiple times and pick the one with the lowest test error.
True or False?

TRUE

3. In two sentences or less, describe the bias variance tradeoff. (1 pt)

Variance is the amount by which $\hat{f}$ changes if estimated using different training data and bias is the error that is introduced when an overly simplistic model is used on a more complet dataset. A tradeoff occurs when, as we use more flexible models that better fit our data, variance increases while bias decreases. 

# Data Questions

```{r}

CAS <- data(CASchools)

#to find number of observations and variables: 


#dim(CAS) 

#to find data type contained w/in each column: 

#str(CAS)

#to find values missing: 

#map(CAS, ~sum(is.na(.)))

#find mean of students: 

#mean(CAS$students)

#find standard deviation of number of computers: 

#sd(CAS$computer)

# How many observations would it drop if you limited the sample to schools with 500+ students?:

#drop <- CAS %>% 
#        filter(students >= 500) %>% 
#        nrow()

```

1. How many observations are in the dataset? (0.5 pts): 420 observations

2. How many variables are in the dataset? (0.5 pts): 14 variables 

3. Are any of the columns categorical? (0.5 pts): Yes, 4 - District (chr), School (chr), County (factor), and Grades (factor). The rest are numeric. 

4. Are any values missing? (0.5 pts): No.

5. What is the mean number of students in the school? (0.5 pts): 2628.79

6. What is the standard deviation of number of computers? (0.5 pts): 441.34

7. What does the calworks variable measure? Hint: Read the codebook! (0.5 pts): "Percent qualifying for CalWorks (income assistance)" 

8. How many observations would it drop if you limited the sample to schools with 500+ students? (0.5 pts): We are left with 281 rows, meaning that we dropped 139 rows.  

### Sort the data by number of students (ascending) and put the first 80 in the test set and
the last 200 in the training set. Include only the following variables in the test/training set:
students, teachers, calworks, lunch, computer, expenditure, income, english, and read. Our
outcome variable is going to be the reading scores.




